---
title: "Compare MCMC vs VB estimates"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    highlight: haddock
    number_sections: yes
    theme: cerulean
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = "hide")
```

<!-- Load modules -->

```{r echo=FALSE, message=FALSE}
# library(data.table)
library(tidyverse)
# library(rstan)
library(logitnorm)
# library(gtools)
```

<!-- Load settings -->

```{r echo=FALSE, message=FALSE}
if (grepl("S34-R31YLVDR",Sys.info()['nodename'])) {
  stop("Fix this")
} else if (grepl("ricard",Sys.info()['nodename'])) {
  source("/Users/ricard/bbreg/settings.R")
  source("/Users/ricard/bbreg/utils.R")
  io$data_dir <- "/Users/ricard/data/betabinomial/diff_residual_disp/synthetic/rep1/"
  io$out_dir <- "/Users/ricard/data/betabinomial/mcmc_vs_vb"
} else{
  stop("Computer not recognised")
}
```

## Define data options
```{r}
N_feat <- 300
N_cpgs <- 15
# N_cells <- c(20, 35, 50, 100, 200, 300)
N_cells <- c(100, 300)
```

## Define model_options
```{r}
use_mcmc <- TRUE
```

## Load simulated data

(TO IMPROVE) Just load replicate 1
```{r }
# dt <- list()
# for (i in 1:length(N_cells)) {
#   file <- paste0(io$data_dir, "data_bbreg_feat", N_feat, "_cells", N_cells[i], "_cpgs", N_cpgs, ".rds")
#   dt[[i]] <- readRDS(file)
# }
```

## Load pre-computed models

```{r}
dt <- list()
for (i in 1:length(N_cells)) {
  file <- paste0(io$data_dir, "bbreg_feat", N_feat, "_cells", N_cells[i], "_cpgs", N_cpgs, "_mcmc", use_mcmc, ".rds")
  dt[[i]] <- readRDS(file)
}
```

## Mean - overdispersion true data
```{r, fig.height=1.7, fig.width=4}
for (i in 1:length(N_cells)) {
  par(mfrow = c(1,2))
  plot_mean_disp_data(mu = dt[[i]]$sim_dt$sim_dt_A$theta_true$mu,
                 gamma = dt[[i]]$sim_dt$sim_dt_A$theta_true$gamma,
                 diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                 is_logit = TRUE, title = paste0("Cells ", N_cells[i]))
  plot_mean_disp_data(mu = dt[[i]]$sim_dt$sim_dt_B$theta_true$mu,
                 gamma = dt[[i]]$sim_dt$sim_dt_B$theta_true$gamma,
                 diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                 is_logit = TRUE, title = paste0("Cells ", N_cells[i]))
}
```


## Mean - overdispersion fitted model
Below we show the estimated mean methylation on the x-axis and the estimated overdispersion on the y-axis. Left facet shows the raw estimated overdispersion estimate, which clearly shows a relationship with mean. The blue line is the basis regression fit to this trend, which allows us to remove the mean-overdispersion trend, as shown on the right facet (residual overdispersion).
```{r, echo=FALSE, fig.width=5.8, fig.height=1.9}
for (i in 1:length(N_cells)) {
  ggA <- plot_mean_disp_fit(fit_obj = dt[[i]]$fit_bbreg_A, 
                            diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx, 
                            title = paste0("Group A - Cells ", N_cells[i]))
  ggB <- plot_mean_disp_fit(fit_obj = dt[[i]]$fit_bbreg_B, 
                            diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx, 
                            title = paste0("Group B - Cells ", N_cells[i]))
  print(cowplot::plot_grid(ggA, ggB))
}
```

## Posterior mean estimates $\mu$
```{r, echo=FALSE, fig.width=7.8, fig.height=2.2}
for (i in 1:length(N_cells)) {
  ggA <- plot_posterior_estimates(fit_obj = dt[[i]]$fit_bbreg_A, sim_dt = dt[[i]]$sim_dt$sim_dt_A,
                                  mode = "mu", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                                  title = paste0("Group A - Cells ", N_cells[i]))
  ggB <- plot_posterior_estimates(fit_obj = dt[[i]]$fit_bbreg_B, sim_dt = dt[[i]]$sim_dt$sim_dt_B,
                                  mode = "mu", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                                  title = paste0("Group B - Cells ", N_cells[i]))
  ggAB <- plot_MA_estimates(fit_obj_A = dt[[i]]$fit_bbreg_A, fit_obj_B = dt[[i]]$fit_bbreg_B,
                            mode = "mu", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                            title = paste0("Groups A and B - Cells ", N_cells[i]))
  print(cowplot::plot_grid(ggA, ggB, ggAB, ncol = 3))
}
```


## Posterior overdispersion estimates $\gamma$
```{r, echo=FALSE, fig.width=7.8, fig.height=2.2}
for (i in 1:length(N_cells)) {
  ggA <- plot_posterior_estimates(fit_obj = dt[[i]]$fit_bbreg_A, sim_dt = dt[[i]]$sim_dt$sim_dt_A,
                                  mode = "gamma", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                                  title = paste0("Group A - Cells ", N_cells[i]))
  ggB <- plot_posterior_estimates(fit_obj = dt[[i]]$fit_bbreg_B, sim_dt = dt[[i]]$sim_dt$sim_dt_B,
                                  mode = "gamma", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                                  title = paste0("Group B - Cells ", N_cells[i]))
  ggAB <- plot_MA_estimates(fit_obj_A = dt[[i]]$fit_bbreg_A, fit_obj_B = dt[[i]]$fit_bbreg_B,
                            mode = "gamma", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                            title = paste0("Groups A and B - Cells ", N_cells[i]))
  print(cowplot::plot_grid(ggA, ggB, ggAB, ncol = 3))
}
```


## Posterior residual overdispersion estimates $\epsilon$
```{r, echo=FALSE, fig.width=7.8, fig.height=2.2}
for (i in 1:length(N_cells)) {
  ggA <- plot_posterior_estimates(fit_obj = dt[[i]]$fit_bbreg_A, sim_dt = dt[[i]]$sim_dt$sim_dt_A,
                                  mode = "epsilon", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                                  title = paste0("Group A - Cells ", N_cells[i]))
  ggB <- plot_posterior_estimates(fit_obj = dt[[i]]$fit_bbreg_B, sim_dt = dt[[i]]$sim_dt$sim_dt_B,
                                  mode = "epsilon", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                                  title = paste0("Group B - Cells ", N_cells[i]))
  ggAB <- plot_MA_estimates(fit_obj_A = dt[[i]]$fit_bbreg_A, fit_obj_B = dt[[i]]$fit_bbreg_B,
                            mode = "epsilon", diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                            title = paste0("Groups A and B - Cells ", N_cells[i]))
  print(cowplot::plot_grid(ggA, ggB, ggAB, ncol = 3))
}
```


# Identifying differentially variable features
## Summary of hits
```{r, results=TRUE, message=TRUE}
diff_analysis <- list()
for (i in 1:length(N_cells)) {
  diff_analysis[[i]] <- differential_test(obj_A = dt[[i]]$fit_bbreg_A, 
                                          obj_B = dt[[i]]$fit_bbreg_B)
}
```

## EFDR threshold grid search plots
```{r, echo=FALSE, fig.width=7.8, fig.height=1.4}
for (i in 1:length(N_cells)) {
  ggM <- plot_efdr_efnr_grid_search(diff_obj = diff_analysis[[i]]$diff_mean_thresh, 
                                    target_efdr = diff_analysis[[i]]$opts$efdr_m,
                                    title = paste0("Diff Mean - Cells ", N_cells[i]))
  ggG <- plot_efdr_efnr_grid_search(diff_obj = diff_analysis[[i]]$diff_disp_thresh, 
                                    target_efdr = diff_analysis[[i]]$opts$efdr_g,
                                    title = paste0("Diff Overdispersion - Cells ", N_cells[i]))
  ggE <- plot_efdr_efnr_grid_search(diff_obj = diff_analysis[[i]]$diff_res_disp_thresh, 
                                    target_efdr = diff_analysis[[i]]$opts$efdr_e,
                                    title = paste0("Diff Residual Overdispersion - Cells ", N_cells[i]))
  
  print(cowplot::plot_grid(ggM, ggG, ggE, ncol = 3))
}
```


## Volcano plots

```{r, echo=FALSE, fig.width=8, fig.height=1.7}
for (i in 1:length(N_cells)) {
  ggM <- plot_volcano(diff_obj = diff_analysis[[i]]$mean_summary, mode = "mu",
                      diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                      evidence_thresh = diff_analysis[[i]]$diff_mean_thresh$evidence_thresh,
                      change_thresh = diff_analysis[[i]]$opts$psi_m,
                      xlab = "Log odds ratio (A/B) change", 
                      ylab = "Posterior evidence probability",
                      title = paste0("Diff Mean - Cells ", N_cells[i]))
  ggG <- plot_volcano(diff_obj = diff_analysis[[i]]$disp_summary, mode = "gamma",
                      diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                      evidence_thresh = diff_analysis[[i]]$diff_disp_thresh$evidence_thresh,
                      change_thresh = diff_analysis[[i]]$opts$psi_g,
                      xlab = "Log odds ratio (A/B) change", 
                      ylab = "Posterior evidence probability",
                      title = paste0("Diff Overdispersion - Cells ", N_cells[i]))
  ggE <- plot_volcano(diff_obj = diff_analysis[[i]]$res_disp_summary, mode = "epsilon",
                      diff_feat_idx = dt[[i]]$sim_dt$diff_disp_features$feature_idx,
                      evidence_thresh = diff_analysis[[i]]$diff_res_disp_thresh$evidence_thresh,
                      change_thresh = diff_analysis[[i]]$opts$psi_e,
                      xlab = "Difference (A - B)", 
                      ylab = "Posterior evidence probability",
                      title = paste0("Diff Residual Overdispersion - Cells ", N_cells[i]))
  print(cowplot::plot_grid(ggM, ggG, ggE, ncol = 3))
}
```


## Sample size effect
Compute confusion matrix
```{r}
conf_matrix <- confusion_matrix_disp(dt = dt, diff_analysis = diff_analysis, N_cells = N_cells)
```
Below we show the performance of the model as we increase the total number of cells

### False Discovery Rate
```{r}
knitr::kable(conf_matrix$fdr, digits = 3)
```

### False Negative Rate
```{r}
knitr::kable(conf_matrix$fnr, digits = 3)
```

### False Positive Rate
```{r}
knitr::kable(conf_matrix$fpr, digits = 3)
```

### True Positive Rate / Recall / Sensitivity
```{r}
knitr::kable(conf_matrix$tpr, digits = 3)
```

### Precision
```{r}
knitr::kable(conf_matrix$precision, digits = 3)
```



