---
title: "Compare MCMC vs VB estimates (in one replicate)"
output: 
  html_notebook: 
    highlight: haddock
    theme: cerulean
    number_sections: true
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = "hide")
```

<!-- Load modules -->

```{r echo=FALSE, message=FALSE}
library(data.table)
library(tidyverse)
library(purrr)
library(logitnorm)
library(ggpubr)
```

<!-- Load settings -->

```{r echo=FALSE, message=FALSE}
if (grepl("S34-R31YLVDR",Sys.info()['nodename'])) {
  stop("Fix this")
} else if (grepl("ricard",Sys.info()['nodename'])) {
  source("/Users/ricard/bbreg/settings.R")
  source("/Users/ricard/bbreg/utils.R")
  io$data_dir <- "/Users/ricard/data/betabinomial/diff_residual_disp/synthetic/rep1/"
  io$out_dir <- "/Users/ricard/data/betabinomial/mcmc_vs_vb"
} else{
  stop("Computer not recognised")
}
```

## Define data options
```{r}
N_feat <- 300
N_cpgs <- 15
# N_cells <- c(20, 35, 50, 100, 200, 300)
# N_cells <- c("50","100","200","300")
N_cells <- c("50","100","300")

models <- c("MCMC","VB")
groups <- c("A","B")
```

## Load pre-computed models

```{r}
models_list <- list()
models_list[["MCMC"]] <- list()
models_list[["VB"]] <- list()

for (i in N_cells) {
  file <- paste0(io$data_dir, "bbreg_feat", N_feat, "_cells", i, "_cpgs", N_cpgs,"_mcmcTRUE.rds")
  models_list[["MCMC"]][[i]] <- readRDS(file)
  
  file <- paste0(io$data_dir, "bbreg_feat", N_feat, "_cells", i, "_cpgs", N_cpgs,"_mcmcFALSE.rds")
  models_list[["VB"]][[i]] <- readRDS(file)
}
```

## Plot mean vs overdispersion for the true data
```{r echo=FALSE, fig.align="center", fig.height=7, fig.width=14}
for (i in N_cells) {
    
    par(mfrow = c(1,2))
    
    plot_mean_disp_data(
      mu = models_list[[1]][[i]]$sim_dt$sim_dt_A$theta_true$mu,
      gamma = models_list[[1]][[i]]$sim_dt$sim_dt_A$theta_true$gamma,
      is_logit = TRUE, 
      title = paste0("Cells: ", i, ", Group: A")
    )
    
    plot_mean_disp_data(
      mu = models_list[[1]][[i]]$sim_dt$sim_dt_B$theta_true$mu,
      gamma = models_list[[1]][[i]]$sim_dt$sim_dt_B$theta_true$gamma,
      is_logit = TRUE, 
      title = paste0("Cells: ", i, ", Group: B")
    )
    
}
```



## Plot mean vs overdispersion for the fitted models

```{r echo=FALSE, fig.align="center", fig.height=7, fig.width=14}
for (i in N_cells) {
  for (j in models) {
    
    ggA <- plot_mean_disp_fit(
      fit_obj = models_list[[j]][[i]]$fit_bbreg_A, 
      title = paste0("Cells: ", i, ", Group: A", ", Model: ",j)
    )
    
    ggB <- plot_mean_disp_fit(
      fit_obj = models_list[[j]][[i]]$fit_bbreg_B, 
      title = paste0("Cells: ", i, ", Group: B", ", Model: ",j)
    )
  
    print(cowplot::plot_grid(ggA, ggB))
  }
}
```

## Compare posterior mean estimates $\mu$

```{r echo=FALSE, fig.align="center", fig.height=7, fig.width=14}
for (i in N_cells) {
  for (j in models) {
  
    ggA <- plot_posterior_estimates(
      fit_obj = models_list[[j]][[i]]$fit_bbreg_A, 
      sim_dt = models_list[[j]][[i]]$sim_dt$sim_dt_A,
      mode = "mu", 
      title = paste0("Cells: ", i, ", Group: A", ", Model: ",j)
    )
    
    ggB <- plot_posterior_estimates(
      fit_obj = models_list[[j]][[i]]$fit_bbreg_B, 
      sim_dt = models_list[[j]][[i]]$sim_dt$sim_dt_B,
      mode = "mu",
      title = paste0("Cells: ", i, ", Group: B", ", Model: ",j)
    )
    
    print(cowplot::plot_grid(ggA, ggB, ncol = 2))
  }
}
```

## Posterior overdispersion estimates $\gamma$

```{r echo=FALSE, fig.align="center", fig.height=7, fig.width=14}
for (i in N_cells) {
  for (j in models) {
  
    ggA <- plot_posterior_estimates(
      fit_obj = models_list[[j]][[i]]$fit_bbreg_A, 
      sim_dt = models_list[[j]][[i]]$sim_dt$sim_dt_A,
      mode = "gamma", 
      title = paste0("Cells: ", i, ", Group: A", ", Model: ",j)
    )
    
    ggB <- plot_posterior_estimates(
      fit_obj = models_list[[j]][[i]]$fit_bbreg_B, 
      sim_dt = models_list[[j]][[i]]$sim_dt$sim_dt_B,
      mode = "gamma",
      title = paste0("Cells: ", i, ", Group: B", ", Model: ",j)
    )
    
    print(cowplot::plot_grid(ggA, ggB, ncol = 2))
  }
}
```


## Posterior residual overdispersion estimates $\epsilon$

```{r echo=FALSE, fig.align="center", fig.height=7, fig.width=14}
for (i in N_cells) {
  for (j in models) {
  
    ggA <- plot_posterior_estimates(
      fit_obj = models_list[[j]][[i]]$fit_bbreg_A, 
      sim_dt = models_list[[j]][[i]]$sim_dt$sim_dt_A,
      mode = "epsilon", 
      title = paste0("Cells: ", i, ", Group: A", ", Model: ",j)
    )
    
    ggB <- plot_posterior_estimates(
      fit_obj = models_list[[j]][[i]]$fit_bbreg_B, 
      sim_dt = models_list[[j]][[i]]$sim_dt$sim_dt_B,
      mode = "epsilon",
      title = paste0("Cells: ", i, ", Group: B", ", Model: ",j)
    )
    
    print(cowplot::plot_grid(ggA, ggB, ncol = 2))
  }
}
```

<!-- Extract summary statistics -->

```{r fetch_point_estimates, echo=FALSE, message=FALSE}
dt <- N_cells %>% map(function(i) {
        models %>% map(function(j) {
          groups %>% map(function(k) {
      
        fit_obj <- models_list[[j]][[i]][[paste0("fit_bbreg_",k)]]
        sim_dt <- models_list[[j]][[i]]$sim_dt[[paste0("sim_dt_",k)]]
        
        dt <- data.table(
          ncells = i,
          model = j,
          group = k,
          feature = 1:length(fit_obj$feature_names),
          
          # epsilon
          epsilon_true = sim_dt$epsilon_true$mu,
          epsilon_mean = matrixStats::colMeans2(rstan::extract(fit_obj$posterior)$epsilon),
          epsilon_sd = apply(rstan::extract(fit_obj$posterior)$epsilon, 2, sd),
        
          # mu
          mu_true = sim_dt$theta_true$mu,
          mu_mean = matrixStats::colMeans2(rstan::extract(fit_obj$posterior)$mu),
          mu_sd = apply(rstan::extract(fit_obj$posterior)$mu, 2, sd),
        
          # gamma
          gamma_true = sim_dt$theta_true$gamma,
          gamma_mean = matrixStats::colMeans2(rstan::extract(fit_obj$posterior)$gamma),
          gamma_sd = apply(rstan::extract(fit_obj$posterior)$gamma, 2, sd)
        )
    }) %>% rbindlist
  }) %>% rbindlist
}) %>% rbindlist %>%
  melt(id.vars=c("ncells","model","group","feature"), variable.factor=FALSE) %>%
  .[,estimate:=strsplit(variable,"_") %>% map_chr(2)] %>%
  .[,variable:=strsplit(variable,"_") %>% map_chr(1)]
```

## Plot MCMC vs VB point estimates

```{r echo=FALSE, fig.align="center", fig.height=7, fig.width=14}
for (i in unique(dt$variable)) {
  
  to.plot <- dt[variable==i & estimate!="true"] %>%
    dcast(ncells+group+estimate+feature~model, value.var="value") %>%
  .[,tmp:=sprintf("Estimate: %s, Ncells: %s",estimate,ncells)]
  
  p <- ggscatter(to.plot, x="MCMC", y="VB", size=0.5, color="group") +
    labs(x="MCMC", y="VB") + ggtitle(i) +
    facet_wrap(~tmp, scales="free") +
    geom_abline(slope=1, intercept=0, linetype="solid") +
    theme(
      plot.title = element_text(hjust = 0.5, size=rel(1.2)),
      axis.text = element_text(size=rel(0.9)),
      legend.position = "none"
    )

  print(p)
}
```

## Plot MCMC vs VB distributions

```{r fetch_distributions, echo=FALSE, message=FALSE}
dt <- N_cells %>% map(function(i) {
        models %>% map(function(j) {
          groups %>% map(function(k) {
      
        fit_obj <- models_list[[j]][[i]][[paste0("fit_bbreg_",k)]]
        sim_dt <- models_list[[j]][[i]]$sim_dt[[paste0("sim_dt_",k)]]
        feature_names <- fit_obj$feature_names
        
        dt <- data.table(
          ncells = i,
          model = j,
          group = k,
          feature = feature_names)
        
        # epsilon
        foo <- rstan::extract(fit_obj$posterior)$epsilon %>% 
          as.data.table %>% 
          setnames(feature_names) %>%
          .[,iteration:=1:.N] %>% .[,parameter:="epsilon"] %>%
          melt(id.vars=c("iteration","parameter"), variable.name="feature")
        dt <- dt %>% merge(foo, by="feature")
        
        return(dt)
    }) %>% rbindlist
  }) %>% rbindlist
}) %>% rbindlist
```

Plot posterior distributions for some randomly selected features.  
Samples are selected from iteration > 2500 (half of the total number of iterations)

```{r echo=FALSE, fig.align="center", fig.height=7, fig.width=14}
min.iter <- 2500
features <- head(unique(dt$feature), n=12)

for (i in unique(dt$parameter)) {
  for (j in N_cells) {
    
    to.plot <- dt[group=="A"] %>%
      .[parameter==i & iteration>=min.iter & feature%in%features & ncells==j] %>%
      .[,ncells:=sprintf("Ncells: %s",ncells)]
    
    p <- ggdensity(to.plot, x="value", y="..density..", fill="model") +
      ggtitle(sprintf("%s, Ncells=%s",i,j)) +
      facet_wrap(~feature, scales="free") +
      theme(
        plot.title = element_text(hjust = 0.5, size=rel(1.2)),
        axis.text = element_text(size=rel(0.6)),
        legend.position = "right"
      )

    print(p)
  }
}
```
